---
title: "Final Report"
toc: true
toc-location: right
execute:
  echo: false
  warning: false
  message: false
---

```{r}
#| label: setup
source(here::here("_common.R"))

rubric <- read_csv(here::here('rubrics', '3-final-report.csv'))
maxPoints <- rubric |>
  filter(rating == "Excellent") |>
  summarise(max = sum(maxPoints)) |>
  pull(max)
```

> **Due**: June 25 by 11:59pm
>
> **Weight**: The final report is part of the 40% final project grade.
>
> **Purpose**: Communicate your data visualization project as a complete, reproducible analysis that tells a clear story using at least three visualizations.
>
> **Assessment**: Your submission will be assessed using the [rubric](#grading-rubric) at the bottom of this page.

## Instructions

Use the [project template](../hw/5-final.html) as the starting point for your report. Your final report must be rendered as a self-contained HTML file.

### Required Sections

Your report should include all of the following:

**1. Introduction**

Briefly introduce your topic and state your research question(s). Motivate why this topic is interesting and important. Describe the format option you chose (same question / different audiences; same data / different questions; interactive dashboard) and explain why it fits your project.

**2. Data**

Describe your data source(s):

- Where did the data come from? Include a URL or citation.
- What are the key variables?
- What pre-processing steps did you take (cleaning, filtering, joining)?

**3. Visualizations**

Include **at least three visualizations** that work together to address your research question. For each visualization:

- Give it a descriptive title
- Describe what the chart shows and why you chose this chart type
- Explain how it connects to your research question and the other charts in your project

If you chose the interactive dashboard option, embed your app (via `<iframe>` or a link) and describe the three visual components.

**4. Interpretation**

Summarize your key findings. What insights do your visualizations reveal? What patterns, trends, or comparisons stand out?

**5. Limitations**

Discuss any limitations of your data or your visualizations. What can't you conclude? What would you do differently with more time or data?

**6. Conclusion**

Provide a takeaway message. What should a reader remember after viewing your project?

**7. References**

Cite your data sources, R packages used, and any visualizations or publications that inspired your work.

### Formatting Requirements

- Use the YAML header from the [project template](../hw/5-final.html); update title, author, and date.
- Set `self-contained: true` so the HTML file includes all images and styles.
- Set `code-fold: true` so code is accessible but not distracting.
- All figures must have descriptive alt text or captions.
- File naming: `5-final.html` and `5-final.qmd`.

## Submission

Submit your project as a link to your rendered HTML report. Your GitHub repository should include:

- `5-final.qmd` — your source file
- `5-final.html` — your rendered report
- `data/` — your data files (if small enough to include in the repo)

Submit assignment in GitHub Classroom by **June 25 at 11:59pm**.

## Grading Rubric {#grading-rubric}

Your final report is worth **`r maxPoints` points**, assessed on the following criteria:

```{r}
make_rubric(rubric)
```
