---
title: "Presentation"
toc: true
toc-location: right
execute:
  echo: false
  warning: false
  message: false
---

```{r}
#| label: setup
source(here::here("_common.R"))

rubric <- read_csv(here::here('rubrics', '4-final-presentation.csv'))
maxPoints <- rubric |>
  filter(rating == "Excellent") |>
  summarise(max = sum(maxPoints)) |>
  pull(max)
```

> **When**: June 25, during class
>
> **Weight**: The presentation is part of the 40% final project grade.
>
> **Purpose**: Communicate your data visualization project to your peers and practice giving and receiving constructive feedback.
>
> **Assessment**: Your presentation will be assessed using the [rubric](#grading-rubric) at the bottom of this page.

## Instructions

Each team will present their final project to the class on **June 25**. All team members must participate in the presentation.

### Length

- **10-12 minutes** of presentation per team
- **3-5 minutes** of Q&A and peer feedback after each presentation

### What to Cover

Your presentation should tell the story of your project from start to finish:

1. **Introduce your topic and research question** — why does it matter? who is the audience?
2. **Describe your data** — where did it come from, what does it contain?
3. **Walk through your visualizations** — present each chart, explain your design choices, and what it reveals
4. **Summarize your findings** — what are the key takeaways?
5. **Reflect on limitations** — what would you do differently?

### Slide Design

- Use **slides** (PowerPoint, Google Slides, Quarto Revealjs, or similar)
- Aim for **one key idea per slide**
- Let your **visualizations do the talking** — avoid bullet-point heavy slides
- Ensure text is large enough to read from the back of the room
- All team members should speak

### Peer Feedback

After each presentation, the class will provide structured feedback. Come prepared to give constructive, specific feedback on:

- Clarity of the research question
- Effectiveness of the visualizations
- Quality of the storytelling
- Design choices

Peer feedback and your reflections on it will be part of your final project grade, (25 out of 140 points), so take it seriously!

## Grading Rubric {#grading-rubric}

Your presentation is worth **`r maxPoints` points**, assessed on the following criteria:

```{r}
make_rubric(rubric)
```
